---
title: "Draft"
format: html
editor: visual
editor_options: 
chunk_output_type: console
---


```{r}

# source("../1_clean_data.R")
 source("../outputs/stream_anly.rds")

```

## Background
Schaefer and colleague investigated how water quality is affected following a hurricane.By analyzing long-term weekly records of stream water chemistry from before and after Hurricane Hugo in 1989, in Puerto Rico's Luquillo Mountains they discovered that nutrient levels, particularly potassium and nitrogen, showed a more significant and lasting impact than canopy gaps. These nutrients increased sharply after the disturbance and remained elevated for several years. Schaefer et al.[(200)](https://doi.org/10.1017/s0266467400001358) 

Here is my attempt to reproducing **Figure 3.** from Schaefer et al.[(2000)](https://doi.org/10.1017/s0266467400001358) 

## Data

To recreate **Figure 3.**, I downloaded the following `.csv` files from the ***Environmental Data Initiative***, available at the following link: [(McDowell and INternational Institute Of Tropical Forest (IITF) 2024)](https://doi.org/10.1017/s0266467400001358)

## Methods
[**Flowchart**](images/workflowimage.png)


## Results
The graph below presents a time-series plot of the 9-week moving average stream water concentration before and after the Hurricane Hugo distribution offering additional contrast among watersheds and ions. Schaefer et al.(2000)
The dotted vertical line represents the date of the hurricane impact. 

```{r}
#|Eecho: False

water_plot <- ggplot(data = stream_anly, aes(x = sample_date, 
                            y = conc_aver)) +

  labs(y = "Average Concentration of Nutrient",
       title = "Concentration in Bisley, Puerto Rico streams before and after Hurrican Hugo",
       x = "Year",
       color = "Basins")+
  geom_line(aes(color = basins)) +
  geom_vline(xintercept = as.Date("1989-09-18"),
             linetype = "dotdash") +
  facet_wrap(ions~ .,
             ncol = 1,
             scales = "free_y")+
  theme_bw()
water_plot
```
